Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-05-17 cordillera-cycle.tex
Write down the main points.

@@

@@

@@

@@

Fig.~\ref{fig:atm} about here.

@@

Figs.~\ref{fig:multirec-timeseries} and ~\ref{fig:multirec-snapshots}.

@@

Figs. ~\ref{fig:hires-timeseries} and ~\ref{fig:hires-snapshots},

@@

@@

Fig~\ref{fig:duration} about here.

@@

Fig ~\ref{fig:deglac} about here.

>>>>

The Cordilleran ice sheet has a long history of explorations and geological
investigations. However, the complexity of the evidence is such that no ice-
sheet scale reconstructions are yet possible. Here we use numerical modelling,
calibrated with field evidence to attempt such a reconstruction.

@@

SSA+SIA, pseudo-plastic sliding, temperature-dependent flow.

@@

Thermal model and deformation model.

@@

Positive degree-day model including spatially and seasonally variable standard
deviation (Fig.~\ref{fig:atm}).

@@

Input climatology from North American Regional Reanalysis according to
sensitivity study from the previous paper (Fig.~\ref{fig:atm}).

@@

All six simulations consistently produce two major glaciations during MIS~4 and
MIS~2 (Figs.~\ref{fig:multirec-timeseries} and~\ref{fig:multirec-snapshots}).
However the timing is much different. Notably the oceanic core forcing result
in much too early last glacial maximum.

@@

We re-run the GRIP (?) and Vostok (?) core simulations at higher resolution
(Figs.~\ref{fig:hires-timeseries} and ~\ref{fig:hires-snapshots}). Results are
consistent with lower resolution but differences appear due to small-scale
topography better capturing mass-balance lapse-rate effects. In general the
Greenland-based simulation show much more variability.

@@

Timing of glaciation is better reproduced by Greenland and Antarctic cores than
nearby oceanic records. This is because the oceanic cores are affected by
variations in strength of a local current (the California current) associated
with the presence of the ice sheet. The timing of the LGM is better reproduced
by Antarctic than Greenland records.

@@

During most of the glacial cycle, ice cover is restricted to high mountain
areas. This is particularly true when using the Greenland-based forcing
(Fig~\ref{fig:duration}). Most of the glacial maxima margin have a transient
character, showing that the ice sheet is far out of balance with climate. There
persist a central ice dome over the Skeena mountains over the entire glacial
period. This ice dome serves as a nucleation centre at the initiation of the
advance phase preceding the last glacial maximum. It is here because of more
widespread precipitation over the Skeena mountains than across other East-West
transects towards south or north of this region. We interpret this as an
orographic precipitation effect resulting from a "window" through the sharp
range of the Coast Mountains at this latitude (Fig?).

@@

Deglaciation happens first in the southern half and then in the northern half
of the ice-covered area (Fig.~\ref{fig:deglac}). Retreat patterns are different
than reconstructed over central BC, however they match Martin's reconstruction
of Liard lowland "unzipping" pretty well. In the GRIP-driven simulation, the
patterns of de-glaciation and Younger Dryas re-advance match with recent
cosmogenic exposure dates.
